---
title: "SickKidsSucrose"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if (!require(tidyverse)) install.packages("tidyverse")
library(tidyverse)
if (!require(lme4)) install.packages("lme4")
library(lme4)
if (!require(Hmisc)) install.packages("Hmisc")
library(Hmisc)
if (!require(corrplot)) install.packages("corrplot")
library(corrplot)
if (!require(PerformanceAnalytics)) install.packages("PerformanceAnalytics")
library(PerformanceAnalytics)
if (!require(emmeans)) install.packages("emmeans")
library(emmeans)
if (!require(broom.mixed)) install.packages("broom.mixed")
library(broom.mixed)
#if (!require(performance)) install.packages("performance")
#library(performance)
```


# Read Data

```{r}
fulldata = read.csv("sucrose_master_data2.csv")
```

# Subjects with full runs

```{r}
fullruns <- fulldata %>% drop_na()
fullruns$AGE <- as.factor(fullruns$AGE)
head(fullruns)
```

# Histograms

```{r}
hist(fullruns$gm_m)
```
```{r}
hist(fullruns$Sucrose)
```
```{r}
fullruns$SucroseCat <- cut(fullruns$Sucrose, c(-1,1,30))
summary(fullruns$SucroseCat)
```

# Basic linear model
## Scan Age

```{r}
basic.lm_scanage <- lm(gm_m ~ Scan.Age, data = fullruns)
summary(basic.lm_scanage)
```
### Plot

```{r}
(prelim_plot_scanage <- ggplot(fullruns, aes(x = Scan.Age, y = gm_m)) +
  geom_point() +
  geom_smooth(method = "lm"))
```
### Pearson Correlation
```{r}
cor(fullruns$Scan.Age, fullruns$gm_m, method="pearson")
cor.test(fullruns$Scan.Age, fullruns$gm_m, method="pearson")
```
### Residuals
```{r}
plot(basic.lm_scanage, which = 1)
```
### Q-Q Plot
```{r}
plot(basic.lm_scanage, which = 2)
```

## Birth.Age
```{r}
basic.lm_birthage <- lm(gm_m ~ Birth.age, data = fullruns)
summary(basic.lm_birthage)
```

### Plot
```{r}
(prelim_plot_birthage <- ggplot(fullruns, aes(x = Birth.age, y = gm_m)) +
  geom_point() +
  geom_smooth(method = "lm"))
```

## FD

```{r}
basic.lm_fd <- lm(gm_m ~ FD, data = fullruns)
summary(basic.lm_fd)
```
### Plot
```{r}
(prelim_plot_fd <- ggplot(fullruns, aes(x = FD, y = gm_m)) +
  geom_point() +
  geom_smooth(method = "lm"))
```
### Pearson Correlation
```{r}
cor(fullruns$FD, fullruns$gm_m, method="pearson")
cor.test(fullruns$FD, fullruns$gm_m, method="pearson")
```
### FD vs Age

```{r}
basic.lm_fdage <- lm(FD ~ Scan.Age, data = fullruns)
summary(basic.lm_fdage)
```
### Plot
```{r}
(prelim_plot_fdage <- ggplot(fullruns, aes(x = Scan.Age, y = FD)) +
  geom_point() +
  geom_smooth(method = "lm"))
```
### Pearson Correlation
```{r}
cor(fullruns$FD, fullruns$Scan.Age, method="pearson")
cor.test(fullruns$FD, fullruns$Scan.Age, method="pearson")
```


## Sex
```{r}
boxplot(gm_m ~ sex, data = fullruns)
```
## Sucrose

```{r}
basic.lm_sucrose <- lm(gm_m ~ Sucrose, data = fullruns)
summary(basic.lm_sucrose)
```
### Plot

```{r}
(prelim_plot_sucrose <- ggplot(fullruns, aes(x = Sucrose, y = gm_m)) +
  geom_point() +
  geom_smooth(method = "lm"))
```
### Residuals
```{r}
plot(basic.lm_sucrose, which = 1)
```
### Q-Q Plot
```{r}
plot(basic.lm_sucrose, which = 2)
```
### Pearson Correlation
```{r}
cor(fullruns$gm_m, fullruns$Sucrose, method="pearson")
cor.test(fullruns$gm_m, fullruns$Sucrose, method="pearson")
```

## H vs Age vs SucroseCat

```{r}
(prelim_plot_sucrose <- ggplot(fullruns, aes(x = Scan.Age, y = gm_m, colour=SucroseCat)) +
  geom_point() +
  geom_smooth(method = "lm"))
```
```{r}
lowsucrose <- subset(fullruns, Sucrose < 1)
highsucrose <- subset(fullruns, Sucrose >= 1)
```
```{r}
install.packages("remotes")
remotes::install_github("GRousselet/bootcorci")
```
```{r}
library(bootcorci)
bootstrapresults <- twocorci(lowsucrose$Scan.Age,lowsucrose$gm_m,highsucrose$Scan.Age,highsucrose$gm_m)
bootstrapresults$estimate1
bootstrapresults$estimate2
bootstrapresults$difference
bootstrapresults$conf.int
bootstrapresults$p.value
```

# Correlation Matrix
```{r}
correlationdata <- fullruns[,c(3,4,5,14,15,17,18,20,31)]
head(correlationdata)
```
## Correlation Matrix
```{r}
res <- cor(correlationdata)
round(res, 2)
```
## Coefficients and p-Values
```{r}
res2 <- rcorr(as.matrix(correlationdata))
res2
```
```{r}
# ++++++++++++++++++++++++++++
# flattenCorrMatrix
# ++++++++++++++++++++++++++++
# cormat : matrix of the correlation coefficients
# pmat : matrix of the correlation p-values
flattenCorrMatrix <- function(cormat, pmat) {
  ut <- upper.tri(cormat)
  data.frame(
    row = rownames(cormat)[row(cormat)[ut]],
    column = rownames(cormat)[col(cormat)[ut]],
    cor  =(cormat)[ut],
    p = pmat[ut]
    )
}
```

```{r}
flattenCorrMatrix(res2$r, res2$P)
```
## Plot
```{r}
corrplot(res, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45)
```
## Insignificant correlations are left blank
```{r}
#corrplot(res2$r, type="upper", order="hclust", 
#         p.mat = res2$P, sig.level = 0.01, insig = "blank")
```

## Chart Correlation
```{r}
chart.Correlation(correlationdata, histogram=TRUE, pch=19)
```

# Mixed Linear Effects Model (continuous Age)
```{r}
mixed.lmer <- lmer(gm_m ~ Scan.Age + FD + (1|Subject.ID), data = fullruns)
summary(mixed.lmer)
```
## Plot
```{r}
plot(mixed.lmer)
```
```{r}
qqnorm(resid(mixed.lmer))
qqline(resid(mixed.lmer))
```


# Mixed Linear Effects Model (categorical Age)

```{r}
mixed.lmer_cA <- lmer(gm_m ~ AGE + FD + (1|Subject.ID), REML=FALSE, data = fullruns)
summary(mixed.lmer)
```
```{r}
mixed.lmer_cA.null <- lmer(gm_m ~ FD + (1|Subject.ID), REML=FALSE, data = fullruns)
anova(mixed.lmer_cA.null,mixed.lmer_cA)
```
```{r}
boxplot(gm_m ~ AGE, data = fullruns)
```

```{r}
emm.cA <- emmeans(mixed.lmer_cA, ~AGE)
output.cA <- contrast(emm.cA, method = "pairwise")
tidy.cA <- tidy(output.cA, conf.int = T) %>% 
  select(contrast, estimate, p.value, conf.low, conf.high)
tidy.cA$p.adj <- p.adjust(tidy.cA$p.value, method = "holm")
tidy.cA
```

# Over all ROIs H
```{r}
roidata <- fullruns %>% pivot_longer(ends_with('_m'), names_to = 'ROI', values_to = 'H')
roidata$ROI <- as.factor(roidata$ROI)
head(roidata)
```

```{r}
ggplot(aes(AGE, H), data = roidata) + 
  geom_boxplot() + 
  facet_wrap(~ ROI) 
```

```{r}
roi.model <- lmer(H ~ AGE*ROI + FD + (1|Subject.ID), REML=FALSE, data = roidata)
roi.null <- lmer(H ~ ROI + FD + (1|Subject.ID), REML=FALSE, data = roidata)
anova(roi.model, roi.null)

emm.roi <- emmeans(roi.model, ~AGE + ROI)
output.roi <- contrast(emm.roi, method = "pairwise", by = "ROI")
tidy.roi <- tidy(output.roi, conf.int = T) %>% 
  select(ROI, contrast, estimate, p.value, conf.low, conf.high)
tidy.roi$p.adj <- p.adjust(tidy.roi$p.value, method = "holm")
tidy.roi
```

# Over all ROIs STD
```{r}
roidata_sd <- fullruns %>% pivot_longer(ends_with('_std'), names_to = 'ROI', values_to = 'SD')
roidata_sd$ROI <- as.factor(roidata_sd$ROI)
head(roidata_sd)
```

```{r}
ggplot(aes(AGE, SD), data = roidata_sd) + 
  geom_boxplot() + 
  facet_wrap(~ ROI) 
```

```{r}
roi_sd.model <- lmer(SD ~ AGE*ROI + FD + (1|Subject.ID), REML=FALSE, data = roidata_sd)
roi_sd.null <- lmer(SD ~ ROI + FD + (1|Subject.ID), REML=FALSE, data = roidata_sd)
anova(roi_sd.model, roi_sd.null)

emm_sd.roi <- emmeans(roi_sd.model, ~AGE + ROI)
output_sd.roi <- contrast(emm_sd.roi, method = "pairwise", by = "ROI")
tidy_sd.roi <- tidy(output_sd.roi, conf.int = T) %>% 
  select(ROI, contrast, estimate, p.value, conf.low, conf.high)
tidy_sd.roi$p.adj <- p.adjust(tidy_sd.roi$p.value, method = "holm")
tidy_sd.roi
```